Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-01-08 paper/paper.tex
revision of discussion and acknowledgement

\section{Discussion and conclusion}
In this paper, we introduce an abstract algebra formalism to compute properties of independent sets.
The properties include the MIS size, number of independent set of a given size, and enumeration of independent sets of a give size.
For each property, we design a algebra being a commutative semiring, and map the property computation to the contraction of a tensor network with such element type.
The data types introduced in the main text can be diagramatically represented as the following, where we use overlap to represent two algebras can be combined to create a new algebra.

@@

%We derived the backward rule for tropical tensor network to bound the search of solution space.
%Although many of these properties are global, we can encode them to different tensor element types as commutative semirings.
Except for real numbers, we have polynomials and truncated polynomials algebras. We combine them with set algebra and bit string algebra for set enumeration and sampling.
We call a tensor network with generic tensor element type a generic tensor network, and its power is not limited to solving the independent set properties.
In \App{app:otherproblems}, we show how to map the maximal independent set problem, matching problem, $k$-colouring problem, max cut problem and set packing problem to tensor networks.
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
%Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
%We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
We show some of the Julia language implementations in Appendix~\ref{sec:technical} and you will find it surprisingly short.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}.
%What we need to do is just defining two operations $\oplus$ and $\odot$ and two special elements $\mymathbb{0}$ and $\mymathbb{1}$.
%The style that we program is called generic programming,

\section*{Acknowledgments}
We would like to thank Pan Zhang for sharing his python code for optimizing contraction orders of a tensor network.
We would like to acknowledge Sepehr Ebadi, Maddie Cain and Leo Zhou for popping up inspiring questions about independent sets,
their questions are driving forces of this project.
Thank Chris Elord for helping us writing the fastest matrix multiplication libary for tropical numbers, TropicalGEMM.jl, he is a man of speed!
Thank other open source software developers: Roger Luo, Time Besard and Katharine Hyatt for solving issues voluntarily.
\blue{funding information}

>>>>

\section{Discussion and conclusion}
In this paper, we introduce a method to use generic programming tensor networks to compute a number of different independent set properties, including the size of the MISs and the number and enumeration of independent sets of a given size. For each property, we design an algebra of a commutative semiring with a particular element type and map the computation of the property to the contraction of a tensor network with such an element type. The data types introduced in the main text can be represented in the following diagram, where we use overlap to represent two algebras that can be combined to create a new algebra.

@@

Other than the the case with real numbers, we have polynomials and truncated polynomials as algebra types. We combine them with the set algebra and the bitstring algebra for set enumeration and sampling. 
This approach to use generic tensor networks can be easily extended to compute properties of maximal independent sets and a variety of other combinatorial problems such as the matching problem, the k-coloring problem, the max-cut problem, and the set packing problem, as detailed in \App{app:otherproblems}. 
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
%Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
%We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
By adapting to the style of generic programming, we can implement all the algorithms without much effort.
We show some of the Julia language implementations in Appendix~\ref{sec:technical}.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}. 
We expect our tool can be used to understand and study many interesting applications of independent sets and beyond. This approach of generic tensor networks should also inspire future works on tensor network computations.

\section*{Acknowledgments}
We would like to thank Pan Zhang for sharing his python code for optimizing contraction orders of a tensor network.
We acknowledge Sepehr Ebadi, Maddie Cain, and Leo Zhou for coming up with many interesting questions about independent sets and their questions strongly motivated the development of this project.
We thank Chris Elord for helping us to write the fastest matrix multiplication library for tropical numbers, TropicalGEMM.jl. % he is a man of speed!
We would also like to thank a number of open-source software developers, including Roger Luo, Time Besard, and Katharine Hyatt, for solving some implementation issues voluntarily.
\blue{funding information}
